\documentclass[11pt]{article}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{placeins}
\title{Professional Issues}
\author{Kevin Macharia(I56/87243/2017) \\ Assignment Two }
\date{\today}
\begin{document}
\maketitle
\paragraph{Question:}Study the following problem statement carefully and answer the questions that
follow: In Kiambu peri-urban area, there is stiff competition for milk between
the formal and informal market outlets, yet many smallholder milk producers
sell milk at low prices, not competitive enough to ensure positive returns to
external inputs they use to support intensive milk production. It is suggested
that market information flow between milk producers and market actors impact
on the price that farmers receive for their milk, and access and use of market
information could explain why some are able to sell competitively while others
are not. Farmers may not be seeking for market information among the market
outlets, if they do, there could be information barriers in access, timeliness and
reliability. Knowledge of the flow and use of market information on milk prices
would be useful in designing effective price information dissemination strategies
to help farmers sell their milk at profitable prices and realize positive returns to
their investment for improved income and food security from dairy production. \\
Answer the following questions in relation to this problem statement:
\begin{enumerate}
\item \textbf{Formulate three research objectives.}
  \begin{itemize}
  \item To determine if the market infomation received by farmers in Kiambu area is reliable and timely.
  \item To identify where in the flow of information are greater barriers and the type of barrier experienced by small scale milk producers and the actors in the formal and informal market actors.
  \item To establish whether small scale milk producers seek for market information about which market outlets offer competitive prices.
  \end{itemize}
\item \textbf{State three research questions.}
  \begin{itemize}
  \item Which communication channel do you use to know the market price of milk?
  \item How reliable and timely is the information on market milk price communication from market actors for small scale milk producers?
  \item Is sourcing of market milk price associated with sale of milk at competitive price?
  \end{itemize}
\item \textbf{State three research hypothesis.}
  \begin{itemize}
  \item \textbf{Hypothesis Ho:} Milk producers that source market information are able
to sell their milk at competitive prices than those not sourcing for market information. \\
\textbf{Hypothesis Ha:} Milk producers that source market information are not able to sell their milk at competitive prices compared to those not sourcing for market information.
  \item \textbf{Hypothesis Ho:} The market information on milk price from market actors for
    small scale milk producers is reliable and timely. \\
    \textbf{Hypothesis Ha:} The market information on milk price from market actors for
small scale milk producers is neither reliable nor timely.
    
  \item \textbf{Hypothesis Ho:} The frequency of using channels of information delivery are
    greatly correlated with reliability, accuracy, and timelines of relaying the milk price information. \\
    \textbf{Hypothesis Ha:} The frequency of using channels of information delivery is not correlated with reliability, accuracy, and timelines of relaying the milk price information.
  \end{itemize}
\item \textbf{Briefly describe the suitable statistical design the researcher will
  use to achieve the stated objectives.} \\
   The best statistical design would be the observational study and specifically
a retrospective study which would observe the small scale milk producers popu-
lation in the present by using a sample survey and collect information from the
sample about how the market information has already affected the milk prices.
\item \textbf{What are the suitable measurements variables from your stated objectives?} \\
They are milk prices, quantity of milk sold, market information perception and intermediary supplier.
\item \textbf{Advise the researchers on the suitable statistical methods that can
be used in analyzing the type of data expected to be collected from
your stated objectives} \\
The researcher can use R to perform statistical analyses. Paired t-test,
factor analysis, ANOVA, and linear regression prove to suit the above scenario.
They might also have to test correlation of the data collected as well
as the statistical significance of the test results (p-value).
\item \textbf{Provide hypothetical tables and figures expected to be generated from the analysis of collected data.} \\
  \begin{table}[htbp]
    \centering
    \begin{tabular}{|l|l|l|l|l|}
      \hline
      \multicolumn{5}{|c|}{Factor correlations} \\[2mm] \hline
      Factor&Factor1&Factor2&Factor3&Factor4\\ [4mm] \hline
      1.Milk prices&1.0&0.6&0.00&0.00 \\[4mm] \hline
      2.Milk quantity&0.6&1.0&0.00&0.00 \\[4mm]\hline
      3.No. of households&0.00&0.00&1.0&0.2 \\[4mm]\hline
      4.Intermediary supplier&0.00&0.00&0.2&1.0 \\[4mm]\hline
    \end{tabular}
    \caption{Correlations table}
    \label{table:correlation}
  \end{table}
  \begin{table}[t!]
    \centering
    \begin{tabular}{|l|l|l|l|l|l|}
      \hline
      \multicolumn{6}{|c|}{ANOVA} \\[2mm] \hline
      ANOVA&df&SS&MSS&F&Significance of F\\ [4mm] \hline
      Regression&1&1300&1300&34.67 & .000160053 \\[4mm] \hline
      Residual&12&450&37.5&0.00&0.00 \\[4mm]\hline
      Total&13&1750&134.6&0.0&0.00 \\[4mm]\hline
    \end{tabular}
    \caption{Anova table}
    \label{table:anova}
    \vspace{1cm}
    \begin{tabular}{|l|l|}
      \hline
      \multicolumn{2}{|c|}{Regression} \\[2mm] \hline
      1&Regression\\ [4mm] \hline
      Multiple R&0.56743826 \\[4mm] \hline
      R Square&0.87234567 \\[4mm]\hline
      Adjusted R Square&0.76214568 \\[4mm]\hline
      Standard Error&4.67156702 \\[4mm]\hline
      Observations&13 \\[4mm]\hline
    \end{tabular}
    \caption{Regression table}
    \label{table:regression}
  \end{table}
      \FloatBarrier
  The reseracher will use correlation to test if there are relationship between the variables, ANOVA and Regression to test the individual significance of each and /or individual variables.
\end{enumerate}

\end{document}
